### 2018-11-25 chapters/5-acknowledgements.tex

\section{Acknowledgements}

This study was financially supported by grants from Instrumentarium Research
Foundation, Sigrid Jusélius Foundation, Turku University Hospital, TYKS-SAPA
research fund, Finnish Cancer Society, Finnish Cultural Foundation, and Orion
Research Foundation. PT was supported by a Clinical Researcher Funding from the
Academy of Finland. We thank Jaakko Liippo (Turku University Hospital, Turku,
Finland) for his help in scanning the histological slides, and Anitta Entonen
(Turku University Hospital, Turku, Finland) for her help with patient enrollment
and assistance during MRI examinations.