### 2019-03-07 response_2019-03.txt
added rebuttal letter and draft for a response

Draft of reply to reviewers. Original comments are included, responses begin
with "*".

Reviewer #1:

In the manuscript "Machine learning of prostate MRI: towards improved
non-invasive prostate cancer characterization", Toivonen et al. develop and
evaluate a radiomics-based classifier system for prediction of prostate cancer
Gleason score. The authors use radiomics features from T2-weighted imaging
(T2w), diffusion weighted imaging (DWI), and T2-mapping (T2) from patients that
subsequently underwent prostatectomy. The DWI data sets were post-processed
using monoexponential and kurtosis models.

The paper is well written, and the approach is detailed sufficiently. As a
general comment on the recent surge of published radiomics papers, while each
paper identifies significant radiomics features, no effort has been made to
connect any of these features with previously published ones. This manuscript is
no exception. More than 1000 features are extracted from four types of images.
There are long lists of variables for each of the four image types; it is not
clear what will be the impact of these findings for future applications. In
addition, the data are acquired in a single institution with the same magnet and
imaging sequences. Again, it is not clear how applicable the identified features
will be in other circumstances. The models are not validated in an independent
dataset.

* [What do they mean by "connecting the features with previously published
  ones"? Comparing them to existing literature? This was done a little, could be
  done more and with more recent papers.]
* TODO: Update and expand literature survey and comparison.
* TODO: Mention in Discussion as a drawback, that there was no testing with
  independent data set (although the machine learning method used does not
  require a separate testing set).

The authors analyze a very rich and unique dataset. For instance, the DWI data
contains 12 b-values. In my view, there are missed opportunities to address
important questions, such as: is there an added value for mapping T2; which
three b-values result in apparent diffusion coefficient (ADC) with the highest
correlation with Gleason Score.

* It is mentioned in the Discussion section, line 556, that adding texture
  features from T2 did not improve classifier performance.
* The evaluation of b-values was not included in the goals of this study, as it
  can be addressed separately.

Figure 1 is misleading; there is little correspondence between the figure legend
and the content. The Field of View, displayed on T2-w and DWI are different. The
labels of the radiomics features are shifted.

* TODO: Clarify Fig 1 and make cosmetic fixes.

Figure 2: Please, list the abbreviations. Why the RP lesions appear differently
on the different sequences?

* TODO: Open abbreviations in Fig 2.
* The lesions look a bit different in the histology and the different sequences,
  because they did not include slices from exactly same locations.

Reviewer #2:

The manuscript showed that a machine learning-based model on imaging features
from multiple MRI sequences provides improved predictive power to discriminate
tumors with 3+3 vs >3+3 Gleason scores in prostate cancer. This paper was well
written and authors performed several tests to show the superiority of the
proposed method. I would suggest to accept this paper if authors can address the
following issues.

1. It is not clear why authors extracted imaging features in 2D slices. Please
explain why you did use 3D volume images.

* We extracted texture features from 2D slices, because the images were highly
anisotropic (about 5x1x1 mm³). This is mentioned in Section 2.5, line 185.
Various slices from 3D images were used in order to make use of more data.

2. Shape-based features were not used. Do you think that shape-based features
are not relevant to Gleason scores. Please mention this in the paper.

* Including shape-based features might be useful for Gleason score
characterization. However, it would require a considerable amount of additional
work. In this manuscript we decided to focus on texture features.
* TODO: Mention how shape features have been used in existing literature.

3. Is there any correlation in imaging features between different image types?

* TODO: How should we test feature correlation between image types?

4. Table 5 shows the improved performance when imaging features from different
image types were combined. The models including all features are not practical
in terms of interpretation and model building. What are the results if 18
statistical features and top 1% features were combined.

* TODO: Estimate classification performance with a combination of statistical
  and top 1% (5%?) texture features.